---
output:
  html_document: default
  pdf_document: default
---

/***********************************************
** File: Assignment2-PartB.sql
** Desc: Manipulating, Categorizing, Sorting and Grouping & Summarizing Data
** Author: Jianghong Man
** Date: 10/30/2020
************************************************/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Initialize libraries

```{r}
library(RMySQL);
library(ggplot2);
library(DBI);
```

Set up the connection

```{r}
connection = dbConnect(MySQL(),user="root", password="rootroot", 
                dbname="sakila", host="localhost");
```

# QUESTION 1

1a) Show the list of databases.

```{r}
myQuery1a <-
"show databases;"
dbGetQuery(connection, myQuery1a)
```

1b) Select sakila database.

```{r}
myQuery1b <-
"USE sakila; "
dbGetQuery(connection, myQuery1b)
```

1c) Show all tables in the sakila database.

```{r}
myQuery1c <- 
"SHOW FULL TABLES;"
c1 <- dbGetQuery(connection, myQuery1c)
head(c1,10)
```

1d) Show each of the columns along with their data types for the actor table.

```{r}
myQuery1d <- 
"show columns from actor;"
dbGetQuery(connection, myQuery1d)
```

1e) Show the total number of records in the actor table.

```{r}
myQuery1e <- 
"SELECT 
    COUNT(*)
FROM
    actor;"
dbGetQuery(connection, myQuery1e)
```

1f) What is the first name and last name of all the actors in the actor table ?

```{r}
myQuery1f <- 
"SELECT 
    first_name, last_name
FROM
    actor;"
f1 <- dbGetQuery(connection, myQuery1f)
head(f1,10)
```

1g) Insert your first name and middle initial ( in the last name column ) into the actors table.

```{r}
myQuery1g <- 
"
insert into actor (first_name, last_name)  values ('JIANGHONG','JM');"
dbGetQuery(connection, myQuery1g)
```

1h) Update your middle initial with your last name in the actors table.

```{r}
myQuery1h <- 
"
UPDATE actor 
SET 
    last_name = 'MAN'
WHERE
    actor_id = 201; "
dbGetQuery(connection, myQuery1h)
```

1i) Delete the record from the actor table where the first name matches your first name.

Assume that nobody in the database has the same first name as mine.

```{r}
myQuery1i <- 
"
DELETE FROM actor 
where first_name = 'JIANGHONG';"
dbGetQuery(connection, myQuery1i)
```

1j) Create a table payment_type with the following specifications and appropriate data types

```{r}
myQuery1j <- 
"
CREATE TABLE payment_type (     
    payment_type_id smallint,
    type varchar(45)            
);     
"

dbGetQuery(connection, myQuery1j);

myQuery1j1 <- 
"
INSERT INTO payment_type VALUES 
(1,'Credit Card'),              
(2,'Cash'),                     
(3,'Paypal'),                   
(4,'Cheque');                   
"

dbGetQuery(connection, myQuery1j1)
```

1k) Rename table payment_type to payment_types.

```{r}
myQuery1k <- 
"ALTER TABLE payment_type
RENAME TO payment_types;"
dbGetQuery(connection, myQuery1k)
```

1l) Drop the table payment_types.

```{r}
myQuery1l <- 
"DROP TABLE payment_types;"
dbGetQuery(connection, myQuery1l)
```

# QUESTION 2

2a) List all the movies ( title & description ) that are rated PG-13 ?

```{r}
myQuery2a <- 
"SELECT 
    title, description
FROM
    film
WHERE
    rating = 'PG-13';"
a2 <- dbGetQuery(connection, myQuery2a)
head(a2,10)
```

2b) List all movies that are either PG OR PG-13 using IN operator ?

```{r}
myQuery2b <- 
"SELECT 
    title, description
FROM
    film
WHERE
    rating IN ('PG' , 'PG-13');"
b2 <- dbGetQuery(connection, myQuery2b)
head(b2,10)
```

2c) Report all payments greater than and equal to 2$ and Less than equal to 7$ ?

Note : write 2 separate queries conditional operator and BETWEEN keyword

1. 

```{r}
myQuery2c1 <- 
"SELECT 
    *
FROM
    payment
WHERE
    amount >= 2 AND amount <= 7;"
c21 <- dbGetQuery(connection, myQuery2c1)
head(c21,10)
```

2. 

```{r}
myQuery2c2 <- 
"SELECT 
    *
FROM
    payment
WHERE
    amount BETWEEN 2 AND 7;"
c22 <- dbGetQuery(connection, myQuery2c2)
head(c22,10)
```

2d) List all addresses that have phone number that contain digits 589, start with 140 or end with 589

Note : write 3 different queries

1. 

```{r}
myQuery2d1 <- 
"SELECT 
    *
FROM
    address
WHERE
    phone LIKE '%589%' or phone LIKE '140%' or phone LIKE '%589';"
dbGetQuery(connection, myQuery2d1)
```

```{r}
myQuery2d2 <- 
"
SELECT 
    *
FROM
    address
WHERE
    phone REGEXP '589|^140';
"
dbGetQuery(connection, myQuery2d2)
```

```{r}
myQuery2d3 <- 
"SELECT 
    *
FROM
    address
WHERE
    phone LIKE '%589%' or phone LIKE'140%'"
dbGetQuery(connection, myQuery2d3)
```

2e) List all staff members ( first name, last name, email ) whose password is NULL ?

```{r}
myQuery2e <- 
"SELECT 
    first_name, last_name, email
FROM
    staff
WHERE
    password IS NULL;"
dbGetQuery(connection, myQuery2e)
```

2f) Select all films that have title names like ZOO and rental duration greater than or equal to 4

```{r}
myQuery2f <- 
"SELECT 
    *
FROM
    film
WHERE
    title LIKE '%Zoo%'
        AND rental_duration >= 4;"
dbGetQuery(connection, myQuery2f)
```

2g) What is the cost of renting the movie ACADEMY DINOSAUR for 2 weeks?

Note : use of column alias

```{r}
myQuery2g <- 
"SELECT 
    (rental_duration * 7 * 2 * rental_rate) AS cost
FROM
    film
WHERE
    title = 'ACADEMY DINOSAUR';"
dbGetQuery(connection, myQuery2g)
```

2h) List all unique districts where the customers, staff, and stores are located

Note : check for NOT NULL values 

```{r}
myQuery2h <- 
"SELECT DISTINCT
    a.district
FROM
    address AS a,
    customer AS c,
    staff AS w,
    store AS s
WHERE
    a.address_id IN (c.address_id , w.address_id, s.address_id);"
h2 <- dbGetQuery(connection, myQuery2h)
head(h2,10)
```

2i) List the top 10 newest customers across all stores 

Assume that "new customers" represents those who are added to the system latest across all stores. In other words, the 'create_data' is the recent for those customers across all stores. 

```{r}
myQuery2i <- 
"SELECT 
    store_id, customer_id, create_date
FROM
    customer
ORDER BY create_date DESC
LIMIT 10;"
dbGetQuery(connection, myQuery2i)
```

# QUESTION 3

3a) Show total number of movies

```{r}
myQuery3a <- 
"
SELECT 
    COUNT(film_id)
FROM
    film;
"
dbGetQuery(connection, myQuery3a)
```

3b) What is the minimum payment received and max payment received across all transactions ?

```{r}
myQuery3b <- 
"
SELECT 
    MIN(amount), MAX(amount)
FROM
    payment;
"
dbGetQuery(connection, myQuery3b)
```

3c) Number of customers that rented movies between Feb-2005 & May-2005 ( based on paymentDate ).

Assume that there might be one customer rented different movies during this priod of time. We only calculate the distinct number of customers.

Also assume that "between Feb-2005 & May-2005" means including the months Feb-2005 and May-2005.

```{r}
myQuery3c <- 
"
SELECT 
    COUNT(DISTINCT customer_id) AS num
FROM
    payment
WHERE
    payment_date LIKE '2005-02%'
        OR payment_date LIKE '2005-03%'
        OR payment_date LIKE '2005-04%'
        OR payment_date LIKE '2005-05%';
"
dbGetQuery(connection, myQuery3c)
```

3d) List all movies where replacement_cost is greater than 15$ or rental_duration is between 6 & 10 days

```{r}
myQuery3d <- 
"
SELECT 
    *
FROM
    film
WHERE
    replacement_cost > 15
        OR rental_duration BETWEEN 6 AND 10;
"
d3 <- dbGetQuery(connection, myQuery3d)
head(d3,10)
```

3e) What is the total amount spent by customers for movies in the year 2005?

```{r}
myQuery3e <- 
"
SELECT 
    SUM(amount) AS total
FROM
    payment
WHERE
    payment_date LIKE '2005%';
"
dbGetQuery(connection, myQuery3e)
```

3f) What is the average replacement cost across all movies ?

```{r}
myQuery3f <- 
"
SELECT 
    AVG(replacement_cost) AS avg_cost
FROM
    film;
"
dbGetQuery(connection, myQuery3f)
```

3g) What is the standard deviation of rental rate across all movies ?

```{r}
myQuery3g <- 
"
SELECT 
    STD(rental_rate) AS std_rate
FROM
    film;
"
dbGetQuery(connection, myQuery3g)
```

3h) What is the midrange of the rental duration for all movies? 

```{r}
myQuery3h <- 
"
SELECT 
    SUM(rental_duration) / COUNT(rental_duration) AS median
FROM
    (SELECT 
        rental_duration
    FROM
        film
    ORDER BY 1
    LIMIT 500 , 2) AS t;
"
dbGetQuery(connection, myQuery3h)
```

# QUESTION 4

4a) Customers sorted by first Name and last name in ascending order.

```{r}
myQuery4a <- 
"
SELECT 
    *
FROM
    customer
ORDER BY first_name , last_name ASC; 
"
a4 <- dbGetQuery(connection, myQuery4a)
head(a4,10)
```

4b) Count of movies that are either G/NC-17/PG-13/PG/R grouped by rating.

```{r}
myQuery4b <- 
"
SELECT 
    rating, COUNT(film_id)
FROM
    film
WHERE
    rating IN ('G' , 'NC-17', 'PG-13', 'PG', 'R')
GROUP BY rating;
"
dbGetQuery(connection, myQuery4b)
```

4c) Number of addresses in each district.

```{r}
myQuery4c <- 
"
SELECT 
    district, COUNT(address_id)
FROM
    address
GROUP BY district;
"
c4 <- dbGetQuery(connection, myQuery4c)
head(c4,10)
```

4d) Find the movies where rental rate is greater than 1$ and order result set by descending order.

```{r}
myQuery4d <- 
"
SELECT 
    film_id, rental_rate
FROM
    film
WHERE
    rental_rate > 1
ORDER BY rental_rate DESC;
"
d4 <- dbGetQuery(connection, myQuery4d)
head(d4,10)
```

4e) Top 2 movies that are rated R with the highest replacement cost?

```{r}
myQuery4e <- 
"
SELECT 
    film_id, replacement_cost, rating
FROM
    film
WHERE
    rating = 'R'
ORDER BY replacement_cost desc 
LIMIT 2;
"
dbGetQuery(connection, myQuery4e)
```

4f) Find the most frequently occurring (mode) rental rate across products.

```{r}
myQuery4f <- 
"
SELECT 
    rental_rate AS mode, COUNT(rental_rate) AS count
FROM
    film
GROUP BY rental_rate
ORDER BY count DESC
LIMIT 1;
"
dbGetQuery(connection, myQuery4f)
```

4g) Find the top 2 movies with movie length greater than 50 mins and which has commentaries as a special features.

Assume that "has commentaries" means that special_features include "commentaries" but can not only include "commentaries".

```{r}
myQuery4g <- 
"
SELECT 
    film_id, length, special_features
FROM
    film
WHERE
    length > 50
        AND special_features LIKE '%commentaries%'
ORDER BY length DESC
LIMIT 2;
"
dbGetQuery(connection, myQuery4g)
```

4h) List the years which has more than 2 movies released. 

```{r}
myQuery4h <- 
"
SELECT 
    release_year, COUNT(DISTINCT film_id)
FROM
    film
GROUP BY release_year
HAVING count(*)>2;
"
dbGetQuery(connection, myQuery4h)
```


















