/***********************************************
** File: Assignment2-PartC.sql
** Desc: Combining Data, Nested Queries, Views and Indexes, Transforming Data
** Author: Jianghong Man
** Date: 10/30/2020
************************************************/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Initialize libraries

```{r}
library(RMySQL);
library(ggplot2);
library(DBI);
```

Set up the connection

```{r}
connection = dbConnect(MySQL(),user="root", password="rootroot", 
                dbname="sakila", host="localhost")
```

# QUESTION 1

1a) List the actors (firstName, lastName) who acted in more then 25 movies

Note: Also show the count of movies against each actor

```{r}
myQuery1a <-
"
SELECT 
    a.first_name, a.last_name, COUNT(*)
FROM
    actor a
        JOIN
    film_actor f ON a.actor_id = f.actor_id
GROUP BY a.first_name , a.last_name
HAVING COUNT(a.first_name) > 25;
"
a1 <- dbGetQuery(connection, myQuery1a)
head(a1,10)
```

1b) List the actors who have worked in the German language movies

Note: Please execute the below SQL before answering this question.

SET SQL_SAFE_UPDATES=0;
UPDATE film SET language_id=6 WHERE title LIKE "%ACADEMY%";

```{r}
myQuery1b1 <-
"
SET SQL_SAFE_UPDATES=0;
"
dbGetQuery(connection, myQuery1b1)
```

```{r}
myQuery1b2 <-
"
UPDATE film SET language_id=6 WHERE title LIKE '%ACADEMY%';
"
dbGetQuery(connection, myQuery1b2)
```

```{r}
myQuery1b3 <-
"
SELECT 
    a.first_name, a.last_name
FROM
    actor a
        JOIN
    film_actor fa ON a.actor_id = fa.actor_id
        JOIN
    film f ON f.film_id = fa.film_id
        JOIN
    language l ON f.language_id = l.language_id
WHERE
    l.name = 'German';
"
b3 <- dbGetQuery(connection, myQuery1b3)
head(b3,10)
```

1c) List the actors who acted in horror movies.

Note: Show the count of movies against each actor in the result set.

```{r}
myQuery1c <-
"
SELECT 
    a.first_name, a.last_name, COUNT(*) 'Number of horror movies'
FROM
    actor a
        JOIN
    film_actor fa ON a.actor_id = fa.actor_id
        JOIN
    film_category fc ON fa.film_id = fc.film_id
        JOIN
    category c ON fc.category_id = c.category_id
WHERE
    c.name = 'Horror'
GROUP BY a.first_name , a.last_name;
"
c1 <- dbGetQuery(connection, myQuery1c)
head(c1,10)
```

1d) List all customers who rented more than 3 horror movies.

```{r}
myQuery1d <-
"
SELECT 
    b.first_name, b.last_name, a.Horror_movie_rented
FROM
    (SELECT 
        a.customer_id, COUNT(*) 'Horror_movie_rented'
    FROM
        (SELECT 
        r.customer_id, r.rental_id, r.inventory_id, i.film_id
    FROM
        rental r
    JOIN inventory i ON r.inventory_id = i.inventory_id) a
    JOIN (SELECT 
        f.film_id, c.name
    FROM
        film_category f
    JOIN category c ON f.category_id = c.category_id) b ON a.film_id = b.film_id
    WHERE
        b.name = 'Horror'
    GROUP BY a.customer_id
    HAVING Horror_movie_rented > 3) a
        JOIN
    customer b ON a.customer_id = b.customer_id;
"
d1 <- dbGetQuery(connection, myQuery1d)
head(d1,10)
```

1e) List all customers who rented the movie which starred SCARLETT BENING

Assume to list distinct customers because same customers could rent the movie starred SCARLETT BENING several times. 

```{r}
myQuery1e <-
"
SELECT 
    distinct c.customer_id, c.first_name, c.last_name
FROM
    film_actor fa
        JOIN
    actor a ON fa.actor_id = a.actor_id
        JOIN
    inventory i ON fa.film_id = i.film_id
        JOIN
    rental r ON r.inventory_id = i.inventory_id
        JOIN
    customer c ON r.customer_id = c.customer_id
WHERE
    a.first_name = 'SCARLETT'
        && a.last_name = 'BENING';
"
e1 <- dbGetQuery(connection, myQuery1e)
head(e1,10)
```

1f) Which customers residing at postal code 62703 rented movies that were Documentaries.

```{r}
myQuery1f <-
"
SELECT 
    c.first_name, c.last_name
FROM
    customer c
        JOIN
    address a ON c.address_id = a.address_id
        JOIN
    rental r ON c.customer_id = r.customer_id
        JOIN
    inventory i ON r.inventory_id = i.inventory_id
        JOIN
    film_category fc ON fc.film_id = i.film_id
        JOIN
    category ca ON fc.category_id = ca.category_id
WHERE
    ca.name = 'Documentary'
        AND a.postal_code = '62703';
"
dbGetQuery(connection, myQuery1f)
```

1g) Find all the addresses where the second address line is not empty (i.e., contains some text), and return these second addresses sorted.

```{r}
myQuery1g <-
"
SELECT 
    address2
FROM
    address
WHERE
    address2 IS NOT NULL
ORDER BY address2;
"
g1 <- dbGetQuery(connection, myQuery1g)
head(g1,10)
```

1h) How many films involve a “Crocodile” and a “Shark” based on film description?

```{r}
myQuery1h <-
"
SELECT 
    COUNT(*)
FROM
    film_text
WHERE
    description LIKE '%Crocodile%'
        AND description LIKE '%Shark%';
"
dbGetQuery(connection, myQuery1h)
```

1i) List the actors who played in a film involving a “Crocodile” and a “Shark”, along with the release year of the movie, sorted by the actors’ last names.

```{r}
myQuery1i <-
"
SELECT 
    first_name, last_name, release_year
FROM
    film f
        INNER JOIN
    film_actor fa ON fa.film_id = f.film_id
        INNER JOIN
    actor a ON a.actor_id = fa.actor_id
WHERE
    description LIKE '%Crocodile%'
        AND description LIKE '%Shark%'
GROUP BY first_name , last_name , release_year
ORDER BY last_name;
"
i1 <- dbGetQuery(connection, myQuery1i)
head(i1,10)
```

1j) Find all the film categories in which there are between 55 and 65 films. Return the
names of categories and the number of films per category, sorted from highest to lowest by the number of films.

```{r}
myQuery1j <-
"
SELECT 
    c.name, COUNT(*) AS count
FROM
    category c
        INNER JOIN
    film_category f ON c.category_id = f.category_id
GROUP BY name
HAVING count BETWEEN 55 AND 65
ORDER BY count DESC;
"
dbGetQuery(connection, myQuery1j)
```

1k) In which of the film categories is the average difference between the film replacement cost and the rental rate larger than 17$?

```{r}
myQuery1k <-
"
SELECT 
    c.name, (AVG(f.replacement_cost) - AVG(f.rental_rate)) as avg_diff
FROM
    film f
        INNER JOIN
    film_category fc ON f.film_id = fc.film_id
        INNER JOIN
    category c ON c.category_id = fc.category_id
GROUP BY fc.category_id
HAVING (AVG(f.replacement_cost) - AVG(f.rental_rate)) > 17;
"
dbGetQuery(connection, myQuery1k)
```

1l) Many DVD stores produce a daily list of overdue rentals so that customers can be
contacted and asked to return their overdue DVDs. To create such a list, search the rental table for films with a return date that is NULL and where the rental date is further in the past than the rental duration specified in the film table. If so, the film is overdue and we should produce the name of the film along with the customer name and phone number

Assume the the rental date is calculated by the difference between the last update in the system and the rental date instead of the current date minus the rental date. 

```{r}
myQuery1l <-
"
SELECT 
    first_name, last_name, phone, title
FROM
    customer a
        JOIN
    address e ON (a.address_id = e.address_id)
        JOIN
    rental b ON (a.customer_id = b.customer_id)
        JOIN
    inventory c ON (b.inventory_id = c.inventory_id)
        JOIN
    film d ON (c.film_id = d.film_id)
WHERE
    return_date IS NULL
        AND DATEDIFF(CAST(b.last_update AS DATE),
            CAST(rental_date AS DATE)) > rental_duration;
"
dbGetQuery(connection, myQuery1l)
```

1m) Find the list of all customers and staff given a store id

Note : use a set operator, do not remove duplicates

```{r}
myQuery1m <-
"
SELECT 
    c.first_name, c.last_name, c.store_id
FROM
    customer c 
UNION ALL SELECT 
    s.first_name, s.last_name, s.store_id
FROM
    staff s
ORDER BY store_id;
"
m1 <- dbGetQuery(connection, myQuery1m)
head(m1,10)
```

# QUESTION 2

2a) List actors and customers whose first name is the same as the first name of the actor
with ID 8.

```{r}
myQuery2a <-
"
SELECT 
    c.first_name 'customer first name',
    c.last_name 'customer last name',
    a.first_name 'actor first name',
    a.last_name 'actor last name'
FROM
    customer c
        JOIN
    actor a ON c.first_name = a.first_name
WHERE
    a.actor_id = 8;
"
dbGetQuery(connection, myQuery2a)
```

2b) List customers and payment amounts, with payments greater than average the
payment amount

```{r}
myQuery2b <-
"
SELECT 
    c.first_name, c.last_name, amount
FROM
    customer c
        JOIN
    payment p ON c.customer_id = p.customer_id
WHERE
    amount > (SELECT 
            AVG(amount)
        FROM
            payment);
"
b2 <- dbGetQuery(connection, myQuery2b)
head(b2,10)
```

2c) List customers who have rented movies atleast once

Note: use IN clause

```{r}
myQuery2c <-
"
SELECT 
    c.first_name, c.last_name
FROM
    customer c
WHERE
    c.customer_id IN (SELECT 
            customer_id
        FROM
            rental);
"
c2 <- dbGetQuery(connection, myQuery2c)
head(c2,10)
```


2d) Find the floor of the maximum, minimum and average payment amount

```{r}
myQuery2d <-
"
SELECT 
    FLOOR(MAX(amount)), FLOOR(MIN(amount)), FLOOR(AVG(amount))
FROM
    payment;
"
dbGetQuery(connection, myQuery2d)
```

# QUESTION 3

3a) Create a view called actors_portfolio which contains information about actors and films ( including titles and category).

```{r}
myQuery3a1 <-
"
 CREATE OR REPLACE VIEW actors_portfolio AS
        select a.first_name, a.last_name, f.title, c.name
        from actor a 
        join film_actor fa ON a.actor_id = fa.actor_id
        join film f on fa.film_id = f.film_id
        join film_category fc on fc.film_id = f.film_id
        join category c ON fc.category_id = c.category_id
        WITH CHECK OPTION;
"
dbGetQuery(connection, myQuery3a1)
```

```{r}
myQuery3a2 <-
"
SELECT * FROM actors_portfolio;
"
a32 <- dbGetQuery(connection, myQuery3a2)
head(a32,10)
```

3b) Describe the structure of the view and query the view to get information on the actor
ADAM GRANT

```{r}
myQuery3b1 <-
"
DESCRIBE actors_portfolio;
"
dbGetQuery(connection, myQuery3b1)
```

```{r}
myQuery3b2 <-
"
SELECT 
    *
FROM
    actors_portfolio
WHERE
    first_name = 'ADAM'
        AND last_name = 'GRANT';
"
b32 <- dbGetQuery(connection, myQuery3b2)
head(b32,10)
```

3c) Insert a new movie titled Data Hero in Sci-Fi Category starring ADAM GRANT

```{r}
myQuery3c1 <-
"
CREATE Table actors_portfolios AS
        select a.first_name, a.last_name, f.title, c.name
        from actor a 
        join film_actor fa ON a.actor_id = fa.actor_id
        join film f on fa.film_id = f.film_id
        join film_category fc on fc.film_id = f.film_id
        join category c ON fc.category_id = c.category_id
        ;
"
dbGetQuery(connection, myQuery3c1)
```

```{r}
myQuery3c2 <-
"
CREATE OR REPLACE VIEW actors_portfolio AS 
SELECT * 
FROM actors_portfolios with check option;
"
dbGetQuery(connection, myQuery3c2)
```

```{r}
myQuery3c3 <-
"
INSERT INTO actors_portfolio (first_name,last_name,title,name) 
VALUES ('ADAM','GRANT','DATA HERO','Sci_Fi');
"
dbGetQuery(connection, myQuery3c3)
```

```{r}
myQuery3c4 <-
"
SELECT * FROM actors_portfolio WHERE first_name='ADAM' and last_name='GRANT' and title='DATA HERO';
"
dbGetQuery(connection, myQuery3c4)
```

```{r}
# myQuery3c0 <-
# "
# drop view actors_portfolio
# "
# dbGetQuery(connection, myQuery3c0)
```

```{r}
# myQuery3c00 <-
# "
# drop table actors_portfolios
# "
# dbGetQuery(connection, myQuery3c00)
```

# QUESTION 4

4a) Extract the street number ( characters 1 through 4 ) from customer addressLine1

```{r}
myQuery4a <-
"
SELECT 
    SUBSTRING_INDEX(address, ' ', 1)
FROM
    address
LIMIT 5;
"
dbGetQuery(connection, myQuery4a)
```

4b) Find out actors whose last name starts with character A, B or C.

```{r}
myQuery4b <-
"
SELECT 
    first_name, last_name
FROM
    actor
WHERE
    last_name REGEXP '^(A|B|C)';
"
b4 <- dbGetQuery(connection, myQuery4b)
head(b4,10)
```

4c) Find film titles that contains exactly 10 characters

```{r}
myQuery4c <-
"
SELECT 
    title
FROM
    film
WHERE
    LENGTH(REPLACE(title, ' ', '')) = 10;
"
c4 <- dbGetQuery(connection, myQuery4c)
head(c4,10)
```

4d) Format a payment_date using the following format e.g "22/1/2016"

```{r}
myQuery4d <-
"
SELECT 
    DATE_FORMAT(payment_date, '%d/%m/%Y')
FROM
    payment;
"
d4 <- dbGetQuery(connection, myQuery4d)
head(d4,10)
```

4e) Find the number of days between two date values rental_date & return_date

```{r}
myQuery4e <-
"
SELECT 
    DATEDIFF(return_date, rental_date)
FROM
    rental;
"
e4 <- dbGetQuery(connection, myQuery4e)
head(e4,10)
```

# QUESTION 5

Provide 5 additional queries and indicate the specific business use cases they address.

5a) List each film and the number of actors who are listed for that film.

```{r}
myQuery5a <-
"
SELECT 
    flm.title, COUNT(*) number_of_actors
FROM
    film flm
        INNER JOIN
    film_actor fim_act ON flm.film_id = fim_act.film_id
GROUP BY flm.title
ORDER BY number_of_actors DESC;
"
a5 <- dbGetQuery(connection, myQuery5a)
head(a5,10)
```

5b) Display all actors who appear in the film 'QUEEN LUKE'

```{r}
myQuery5b <-
"
SELECT 
    first_name, last_name
FROM
    actor
WHERE
    actor_id IN (SELECT 
            actor_id
        FROM
            film_actor
        WHERE
            film_id IN (SELECT 
                    film_id
                FROM
                    film
                WHERE
                    LOWER(title) = LOWER('QUEEN LUKE')));
"
dbGetQuery(connection, myQuery5b)
```

5c) Identify all movies categorized as Children films in order to target all Children movies for a promotion.

```{r}
myQuery5c <-
"
SELECT 
    film_id, title, release_year
FROM
    film
WHERE
    film_id IN (SELECT 
            film_id
        FROM
            film_category
        WHERE
            category_id IN (SELECT 
                    category_id
                FROM
                    category
                WHERE
                    name = 'Children'));
"
c5 <- dbGetQuery(connection, myQuery5c)
head(c5,10)
```

5d) List the top five genres in gross revenue in descending order.

```{r}
myQuery5d <-
"
SELECT 
    cat.name category_name, SUM(IFNULL(pay.amount, 0)) revenue
FROM
    category cat
        JOIN
    film_category flm_cat ON cat.category_id = flm_cat.category_id
        JOIN
    film fil ON flm_cat.film_id = fil.film_id
        JOIN
    inventory inv ON fil.film_id = inv.film_id
        JOIN
    rental ren ON inv.inventory_id = ren.inventory_id
        JOIN
    payment pay ON ren.rental_id = pay.rental_id
GROUP BY cat.name
ORDER BY revenue DESC
LIMIT 5;
"
dbGetQuery(connection, myQuery5d)
```

5e) Display the least frequently rented movies in ascending order.

```{r}
myQuery5e <-
"
SELECT 
    f.film_id, f.title, A.*
FROM
    film f
        JOIN
    (SELECT 
        inv.film_id, COUNT(ren.rental_id) times_rented
    FROM
        rental ren
    JOIN inventory inv ON ren.inventory_id = inv.inventory_id
    GROUP BY inv.film_id) A ON f.film_id = A.film_id
ORDER BY A.times_rented;
"
e5 <- dbGetQuery(connection, myQuery5e)
head(e5,10)
```
