---
title: "ts_final_project"
author: "Jianghong Man"
date: "5/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(fpp)
library(tseries)
library(ggplot2)
library(forecast)
library(TSA)
library(dplyr) 
library(zoo)
```

# 1. Data Overview

### Data exploration (starting from 1898 Dec).

```{r}
GlobalTemperatures <- read.csv("GlobalTemperatures.csv")
data19 <- GlobalTemperatures$LandAverageTemperature[1788:3192]
ts19 <- ts(data19, start = c(1898,12), frequency = 12)
tsdisplay(ts19)
```

Since the data is too tight and hard to differenciate, we move to excel and python to do a more explicit group of plots. 

For the exploration purpose, we decide to display the data from 2000 to 2015 to show the basic trend. 

# 2. EDA

### Load data and plot the time series of Land Average Temperature of the most recent 15 years (2000 Jan to 2015 Dec)

```{r}
data <- GlobalTemperatures$LandAverageTemperature[3001:3192]
data <- ts(data, start = c(2000, 1), frequency = 12)
tsdisplay(data)
```


```{r}
plot(data)
title(main="Monthly Global Temperatures, in C") 
```

```{r}
#Plot airpass data against time
autoplot(data)+ #Add labels and titles
    labs(x="Date",
         y="Global Temperatures",
         title="Global Temperatures, in C, on Log Scale") + 
    geom_point(color="blue") + # Show the data points in blue
    scale_y_log10() # Put y axis on a logarithmic scale
```

```{r}
## function to standardize (z-score)
stand <- function(x) {
    y <- (x - mean(x,na.rm=TRUE))/sd(x,na.rm=TRUE)
    return(y)
}

# Create zoo object for satandardised anomalies:
stand_data <- zoo(stand(data))
plot(stand_data,main="",ylab="Standardized Temp (Z)",xlab="Year")
```

```{r}
data2006 <- GlobalTemperatures$LandAverageTemperature[3074:3085]

data1986 <- GlobalTemperatures$LandAverageTemperature[2834:2845]

data1886 <- GlobalTemperatures$LandAverageTemperature[1634:1645]

data1786 <- GlobalTemperatures$LandAverageTemperature[434:445]

temp2016 <- ts(c(4.169, 4.451, 6.618, 9.576, 12.338, 14.421, 15.212, 14.795, 13.075, 10.441, 7.281, 4.815), start = c(2016, 1), frequency = 12)


data2006 <- ts(data2006, start = c(2006, 1), frequency = 12)
data1986 <- ts(data1986, start = c(1986, 1), frequency = 12)
data1886 <- ts(data1876, start = c(1876, 1), frequency = 12)

td <- ts(c(data2006,data1986,data1886,temp2016),frequency = 12)
tsdisplay(td)
```

```{r}
##Seasonal plot of airline passengers data
ggseasonplot(td,
            year.labels=TRUE)+
  scale_y_log10()+ # Put y axis on a logarithmic scale
  ylab("Global Temperatures")+ # An alternate way to label Y axis
  ggtitle("Seasonal Plot of Global Temperatures, in C") # Add a title
```

```{r}
##Seasonal plot of airline passengers data
ggseasonplot(data,
            year.labels=TRUE)+
  scale_y_log10()+ # Put y axis on a logarithmic scale
  ylab("Global Temperatures")+ # An alternate way to label Y axis
  ggtitle("Seasonal Plot of Global Temperatures, in C") # Add a title
```

```{r}
plot(decompose(data))
```

# 3. Hypothesis test

### Test whether the data is stationay. 

```{r}
kpss.test(data)
```

### Fit Box-Cox transformation: test whether the data need to stabilize. 

```{r}
BoxCox.lambda(data)
```

Transform the data with box-cox (lambda = 1.499917)

```{r}
lambda <- BoxCox.lambda(data)

data %>% BoxCox(lambda = lambda) 
```

### One seasonal differencing + one non-seasonal differencing

```{r}
tsdisplay(diff(diff(data,lag = 12)),main='Monthly Global Temperatures Flow')
```

### Run hypothesis test to check if it's stationary

```{r}
kpss.test(diff(diff(data, lag = 12)))
```

### Check whether there's missing values in the dataset.（missing value)

```{r}
sum(is.na(data))
```

# 4. Spectral Analysis

```{r}
spectrum(data)
```

```{r}
periodogram(data)
```

```{r}
## Zoom in 
temp <- periodogram(data,xlim=c(0.083,0.0835))
```

f = 0.083, this frequency corresponds to a period of 1/0.083 = 12 days. 
