---
title: "Assignment4_JianghongMan"
author: "Jianghong Man"
date: "9/28/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(ggplot2)
library(dplyr)
library(readxl)
```

## Chapter 14
#### Question 2

a) Y’ = 2 * 6 + 9 = 21

b) 14 = 2X + 9, then X = 2.5

#### Question 4

The standard error of the estimate is a measure of the accuracy of predictions. The formula is as follows:

For population:

$\sigma_{est} = \sqrt{\frac{\Sigma(Y-Y')^2}{N}}$

For sample:

$\sigma_{est} = \sqrt{\frac{\Sigma(Y-Y')^2}{N-2}}$

#### Question 6 chongxinsuan 

```{r}
x <- c(2,4,4,5,6)
y <- c(5,6,7,11,12)
res <- cor.test(x, y, method = "pearson")
res
r <- lm(y ~ x)
summary(r)
```


a) r = 0.9092; p = 0.0324. Therefore, r is significantly different from 0 and we can reject the null hypothesis in this case.

b) Slope = 1.9091, intercept = 0.1818. p-value is 0.0324 for the slope, which indicates that it differs significantly from zero.

c) t for confidence interval is 3.7819, and 95 percent confidence interval:

Lower limit = 0.9940671

Upper limit = 0.1360488

#### Question 8 

t = 1.8517, df = 18, p = 0.0805. Thus, it is not significant at 0.05 level.

#### Question 10 

```{r}
pre <- c(59,52,44,51,42,42,41,45,27,63,54,44,50,47,55,49,45,57,46,60,65,64,50,74,59)
post <- c(56,63,55,50,66,48,58,36,13,50,81,56,64,50,63,57,73,63,46,60,47,73,58,85,44)
re <- lm(post ~ pre)
summary(re)
```


Post = 16.1552 + 0.7869 x Pre

Y’ = 0.7869 * 43 + 16.1552 = 49.9919

#### Question 12

b = r * (Sy/Sx) = -0.6 * (3 / 2.5) = -0.72

A = My - b * Mx = 12 - (-0.72) * 10 = 19.2

Thus, Y’ = 19.2 - 0.72X

#### Question 14
 
True

#### Question 16

False

#### Question 18

```{r}
AM_data <- read_excel("angry_moods.xls")
colnames(AM_data) <- c("Gender", "Sports", "Anger_Out", "Anger_In", 
                       "Control_Out", "Control_In", "Anger_Expression")

r2 <- lm(AM_data$Anger_Out ~ AM_data$Control_Out)
summary(r2)
```

a) slpoe = -0.52413
 
b) intercept = 28.49482

c) Yes

d) p = 2.18e-08 < .001, significant

e) standard error = 3.45

#### Question 19

```{r}
sat_data <- read_excel("sat.xls")
colnames(AM_data) <- c("high_GPA", "math_SAT", "verb_SAT", "comp_GPA", 
                       "univ_GPA")

regression_sat <- lm(sat_data$univ_GPA ~ sat_data$high_GPA)
summary(regression_sat)
```

a) slope = 0.67483

b) y-intercept = 1.09682

c) Y' = 2.2 * 0.67483 + 1.09682 = 2.581446

d) Y' = 4.0 * 0.67483 + 1.09682 = 3.79614

## Chapter 16
#### Question 1

The log transformation can be used to make highly skewed distributions less skewed. It can be valuable both for making patterns in the data more interpretable and for making the relationship between variables clearer. 

#### Question 2

1000

#### Question 3

3,4,5

#### Question 4

-2

#### Question 5

2
