---
title: "Week3_JianghongMan"
author: "Jianghong Man"
date: "10/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Import the library

```{r}
dataPath <- "/Users/nimo/Desktop/31007/3/HW3/"
dat <- read.table(paste(dataPath,'Week3_Test_Sample.csv',sep = '/'), header=TRUE)
```

The sample dat has the following format:

dat$x[1] - mean value of normal distribution;
dat$x[2] - standard deviation of normal distribution;
dat$x[3] - intensity of exponential distribution;
datx[4] : datx[503] sample from uniform distribution on [0.1].

### Sample datNorm from normal distribution with mean dat$x[1] and standard deviation dat$x[2]

```{r}
datNorm <- qnorm(dat$x[4:503], dat$x[1], dat$x[2])
```

### Sample datExp from exponential distribution with intensity dat$x[3].

```{r}
datExp <- qexp(dat$x[4:503], rate = dat$x[3])
```

### Collect all results

```{r}
res<-cbind(datNorm=datNorm,datExp=datExp)
```

### Write the result to a file

```{r}
write.table(res, file = paste(dataPath,'result.csv',sep = '/'), row.names = F)
```

